\documentclass[twoside]{umthesis}
%\documentclass{umthesis}
%\documentclass[draft]{umthesis}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[inner=1.5in,outer=1in,top=1in,bottom=1in,headheight=0pt]{geometry}

\usepackage{amsmath}
% font choice
\usepackage[T1]{fontenc}
%\usepackage{txfonts}
\usepackage{fourier}
\SetMathAlphabet{\mathcal}{normal}{OMS}{cmsy}{m}{n}%revert calligraphy math

% \singlespacing
%\doublespacing
% for clearing the page on final draft mode
\newcommand{\thesisclearpage}{\clearpage}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% extra packages and shortcuts
\usepackage{bm}
\usepackage{amssymb}
\usepackage{microtype}
\usepackage[pdftex]{graphicx}
\usepackage[width=.75\textwidth,font=small,labelfont=bf]{caption}
\usepackage{booktabs} % \toprule, \midrule, \bottomrule
\usepackage{verbatim}
% note: rotating must go AFTER graphicx
\usepackage{subfig} % figure 1a, 1b
\usepackage{rotating} % sideways tables
% note: this file is in $(git rev-parse --show-toplevel)/texmf/etc
\input{SRJinclude.tex}
\setSRJthesisfigurepaths

% For use in the introduction
\newcommand{\chaptersynopsis}[1]{%
\subsubsection{Chapter~\ref{#1}: \nameref{#1}}}
% For referencing equations from a previous chapter; defined like this for
% compatibility with ``single-chapter'' compile
%\newcommand{\tagref}[1]{\tag{\ref{#1}}}


% float setup
\renewcommand\floatpagefraction{.70}
\renewcommand\topfraction{.95}
\renewcommand\bottomfraction{.95}
\renewcommand\textfraction{.1}

% Improve list and enum spacing
\renewcommand{\itemhook}{\par\setstretch{1}%
  \setlength{\topsep}{0pt}%
  \setlength{\parskip}{0pt}%
  \setlength{\partopsep}{.5\baselineskip}%
  \setlength{\parsep}{0pt}%
  \setlength{\itemsep}{0pt}%
}
\renewcommand{\enumhook}{\itemhook}

\makeatletter
\renewcommand\section{\@startsection {section}{1}{\z@}%
                                     {-2.5ex  \@minus -.2ex}%
                                     {1.5ex \@minus .2ex}%
                                     {\normalfont\Large\bfseries}}
\renewcommand\subsection{\@startsection{subsection}{2}{\z@}%
                                     {-1.75ex  \@minus -.2ex}%
                                     {1.25ex \@minus .2ex}%
                                     {\normalfont\large\bfseries}}
\renewcommand\subsubsection{\@startsection{subsubsection}{3}{\z@}%
                                     {-1.5ex  \@minus -.2ex}%
                                     {1.25ex \@minus .2ex}%
                                     {\normalfont\normalsize\bfseries}}
\makeatother  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\author{Seth R.~Johnson}
\title{Anisotropic Diffusion Approximations for Time-dependent Particle
Transport\texorpdfstring{\\%
  \emph{DRAFT: \today}}{}
}

\program{Nuclear Engineering and Radiological Sciences}
\degree{Doctor of Philosophy}
\chaircommitteemember{Edward W.~Larsen}{Professor}
\committeemember{Thomas J.~Downar}{Professor}
\committeemember{James P.~Holloway}{Professor}
\committeemember{William R.~Martin}{Professor}
\committeemember{Katsuyo S.~Thornton}{Professor}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\includeonly{%
introduction,%
trtBackground,%
adDerivation,%
aponeDerivation,%
implementation,%
flatland,%
simpleNumericalResults,%
trtNumericalResults,%
conclusion,%
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

% let ``align'' break anywhere
\allowdisplaybreaks
% keep huge spaces out from between paragraphs, especially when using double
% spacing
\setlength{\parskip}{0pt plus 0pt minus 0pt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% FRONT MATTER
\frontmatter

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frontispiece}
\centering
TENET INSANABILE MULTOS SCRIBENDI CACOETHES
\\
{\small \itshape (The incurable itch to write affects many.)}
\end{frontispiece}
%\begin{frontispiece}
% \begin{flushleft}
% Every morning upon awakening, I experience a supreme pleasure: that of being
% Salvador Dal\'i.
% \end{flushleft}
% ---Salvador Dal\'i
%\begin{flushleft}
%  Only the skilled can judge the skilfulness, but that is not the same as judging the value of the result.
%\end{flushleft}
%---C.~S.~Lewis
%\end{frontispiece}
  
%\begin{frontispiece}
%%\begin{flushleft}
%%For in much wisdom is much vexation,\hfill\\
%%\hspace{1.5em}and he who increases knowledge increases sorrow.
%%\end{flushleft}
%%---Eccl.~1:18
%%THIS PAGE INTENTIONALLY LEFT BLANK
%\begin{flushleft}
%\end{frontispiece}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{dedication}
  To my parents, Doug and Carol.
\end{dedication}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{acknowledgments}
%  TBD
%\begin{comment}
  This dissertation would not have been possible without a lifetime of
  edification by others. The foremost of these mentors are my
  parents, who helped my
  interest in science to blossom at an early age \cite{Lew1989}.
  Other strong influences over the years include Kristen Jones, whose
  leadership in high school science bowl stands out particularly;
  Bruce Freeman, who gave invaluable opinions, advice, and support during my
  undergraduate internships; Kevin Clarno, an
  integral force in shaping my interest in reactor analysis and transport
  methods; and the inimitable Todd Urbatsch, whose work and mentoring at LANL
  gave me a solid foundation for my TRT research projects and codes.

  My most recent mentor, Dr. Ed Larsen, has my greatest appreciation and thanks.
  His methodical derivations---as well as his practically limitless patience for
  an upstart like myself---were essential in my graduate education and
  research experiences. The members of my committee also have my grateful
  appreciation for their help in this work.

  In addition to these more formal mentors are my friends over the years who
  have taken something of a mentoring role. Greg Davidson, Jesse Cheatham, and
  Troy Becker transitioned me into graduate school.
  The mentoring and friendship of Shaun and Adrianne Clarke have
  helped me to grow over these years in Michigan. My brother Aaron and sister
  Naomi have remained a source of motivation and endurance.
  Space does not permit me to name the many other friends and family whose
  encouragement
  has been instrumental in helping me struggle through the monumental task of
  writing this thesis.

  I also thank the National Science Foundation for its financial support via
  the Graduate Research Fellowship Program, the Department of Energy for its
  Nuclear Energy University Programs fellowship, and the NERS department for its
  support and its management of the day-to-day necessities.

  Finally, I offer all glory and praise to Jesus Christ, the Creator and my
  Redeemer. Apart from Him, nothing could be possible---especially not this
  dissertation.
%\end{comment}
\end{acknowledgments}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\begin{preface}
%  before reading this, you should know\dots
%\end{preface}

% list of contents, etc
\tableofcontents
\listoftables
\listoffigures
%\listofappendices

% % the optional normal abstract is formatted the same as preface and acknowledgements,
% % and is listed in the table of contents
% \begin{abstract}
% \end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% MAIN MATTER
\mainmatter

% note that chapter markers MUST go inside the ``include''-d file
\include{introduction}
\include{trtBackground}
\include{adDerivation}
\include{aponeDerivation}
\include{implementation}
\include{flatland}
\include{simpleNumericalResults}
\include{trtNumericalResults}
\include{conclusion}

%\appendix
%\include{pytrt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% BIBLIOGRAPHY
\backmatter
\bibliographystyle{ans}
\bibliography{../SRJall}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% OPTIONAL INDEX
%\phantomsection %makes sure it points to the right page
%\addtocontents{toc}{chapter}{Index}
%\printindex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}
